---
title: "Exploring the metafor package for meta-analysis"
author: "Michael Andreae"
fontfamily: pslatex
output:
  pdf_document: default
---

\setlength{\parindent}{0ex}

##### Authors

Michael H. Andreae, M.D., Department of Anesthesiology & Perioperative Medicine, H187, Penn State Health Milton S. Hershey Medical Center, Penn State College of Medicine, Hershey. PA


\setlength{\parindent}{0ex}

```{r, check_install_libraries, echo=FALSE, message=FALSE, eval=TRUE}
# This chunk will check if all requiered libaries are installed 
# and install the missing libraries

# https://cran.r-project.org/web/packages/pacman/vignettes/Introduction_to_pacman.html

## Option 1
if (!require("pacman")) install.packages("pacman", repos='http://cran.us.r-project.org')
pacman::p_load(knitr, ggplot2, tidyverse, metafor, readxl)

## Option 2, 
### if option one causes an error, comment option one and uncomment option 2:

# if (!require("knitr")) install.packages("knitr", repos='http://cran.us.r-project.org')
# if (!require("ggplot2")) install.packages("ggplot2", repos='http://cran.us.r-project.org')
# if (!require("readxl")) install.packages("readxl", repos='http://cran.us.r-project.org')

## Option 3
### if option 1 and option 2 both throw an error, then:

# install the packages knitr, ggplot2, readxl manuallyand 
# set "eval=FALSE"" for this chunk above in the top line


```

```{r, global_options, echo=FALSE}
suppressWarnings(library(knitr))
# set options
opts_chunk$set(fig.width=8, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
eval_everything=FALSE
```

```{r libraries}
library(metafor)

```

# Documentation in Vignettes
An Introduction is presented in vignettes, expecially in the root directory:

*metafor/Conducting Meta-Analyses in R with the metafor.pdf*

```{r Vignettes, eval=FALSE, echo=TRUE}
browseVignettes("metafor")
```

## Reproducing the example

```{r example1}
dat <- escalc(measure = "RR", ai = tpos, bi = tneg, ci = cpos,
              di = cneg, data = dat.bcg, append = TRUE)
kable(dat)

res <- rma(yi, vi, data = dat)
res

forest(res, slab = paste(dat$author, dat$year, sep = ", "),
       xlim = c(-16, 6), at = log(c(0.05, 0.25, 1, 4)), atransf = exp,
       ilab = cbind(dat$tpos, dat$tneg, dat$cpos, dat$cneg),
       ilab.xpos = c(-9.5, -8, -6, -4.5), cex = 0.75)
op <- par(cex = 0.75, font = 2)
text(c(-9.5, -8, -6, -4.5), 15, c("TB+", "TB-", "TB+", "TB-"))
text(c(-8.75, -5.25), 16, c("Vaccinated", "Control"))
text(-16, 15, "Author(s) and Year", pos = 4)
text(6, 15, "Relative Risk [95% CI]", pos = 2)
par(op)
```

# Cochrane review analysis Regional Anesthesia for Persistent Pain

We import and clean the data extracted during the systematic review and save it as a .Rdata file 

```{r import_ACE151_data, eval=FALSE, echo=TRUE}
library(readxl)
ACE151_raw <- read_excel("C:/Users/mandreae/Dropbox/Cochrane/revision 2014/Updated Search 2016/Overview of all included studies.xlsx")
save(ACE151_raw, file="data/ACE151_raw.Rdata")
rm(ACE151_raw)
```


## Import and clean Breast Data

```{r clean_ACE151_Breast, echo=TRUE, eval=FALSE}
library(tidyverse)
load(file="data/ACE151_raw.Rdata")
ACE151_Breast <- ACE151_raw %>% 
  filter(`Surgery Group` == "Breast") %>%
  transmute(author = factor(Author), 
            year = Year,
            regional = factor(`Anesthesia Technique`), 
            endpoint =as.numeric(`Follow Up (months)`),
            ai = as.numeric(`Events (experimental group)`),
            bi = as.numeric(`N (experimental group)`) - ai,
            ci = as.numeric(`Events (control group)`),
            di = as.numeric(`N (control group)`) - ci,
            singleshot = as.factor(`Single Shot or Continuous`))

kable(ACE151_Breast)
save(ACE151_Breast, file = "results/ACE151_Breast.Rdata")
rm(ACE151_raw, ACE151_Breast)
```

## Compute Breast data

Computing OR for breast studies with follow up at 3 months yields the same results as in Revman.

```{r compute_ACE151_Breast}
library(metafor)
library(tidyverse)
load("results/ACE151_Breast.Rdata")

ACE151_Breast <-   filter(ACE151_Breast, endpoint == 3)
ACE151_Breast <-  escalc(measure = "OR", ai = ai, bi = bi, ci = ci,
              di = di, data = ACE151_Breast, append = TRUE)
  

res <- rma(yi, vi, data = ACE151_Breast)
res
forest(res, slab = paste(ACE151_Breast$author, ACE151_Breast$year, sep = ", "),
       xlim = c(-16, 6), at = log(c(0.05, 0.25, 1, 4)), atransf = exp,
       ilab = cbind(ACE151_Breast$ai, ACE151_Breast$bi, ACE151_Breast$ci, ACE151_Breast$di),
       ilab.xpos = c(-9.5, -8, -6, -4.5), cex = 0.75)
op <- par(cex = 0.75, font = 2)
text(c(-9.5, -8, -6, -4.5), 23, c("PPP+", "PPP-", "PPP+", "PPP-"))
text(c(-8.75, -5.25), 24.5, c("Regional", "Control"))
text(-16, 23, "Author and Year", pos = 4)
text(6, 23, "Odds Ratio [95% CI]", pos = 2)
par(op)
```
