---
title: "Exploring the metafor package for meta-analysis"
author: "Michael Andreae"
fontfamily: pslatex
output:
  pdf_document: default
---

\setlength{\parindent}{0ex}

##### Authors

Michael H. Andreae, M.D., Department of Anesthesiology & Perioperative Medicine, H187, Penn State Health Milton S. Hershey Medical Center, Penn State College of Medicine, Hershey. PA


\setlength{\parindent}{0ex}

```{r, check_install_libraries, echo=FALSE, message=FALSE, eval=TRUE}
# This chunk will check if all requiered libaries are installed 
# and install the missing libraries

# https://cran.r-project.org/web/packages/pacman/vignettes/Introduction_to_pacman.html

## Option 1
if (!require("pacman")) install.packages("pacman", repos='http://cran.us.r-project.org')
pacman::p_load(knitr, ggplot2, tidyverse, metafor)

## Option 2, 
### if option one causes an error, comment option one and uncomment option 2:

# if (!require("knitr")) install.packages("knitr", repos='http://cran.us.r-project.org')
# if (!require("ggplot2")) install.packages("ggplot2", repos='http://cran.us.r-project.org')
# if (!require("readxl")) install.packages("readxl", repos='http://cran.us.r-project.org')

## Option 3
### if option 1 and option 2 both throw an error, then:

# install the packages knitr, ggplot2, readxl manuallyand 
# set "eval=FALSE"" for this chunk above in the top line


```

```{r, global_options, echo=FALSE}
suppressWarnings(library(knitr))
# set options
opts_chunk$set(fig.width=8, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
eval_everything=FALSE
```

```{r libraries}
library(metafor)

```

# Documentation in Vignettes
An Introduction is presented in vignettes, expecially in the root directory:

*metafor/Conducting Meta-Analyses in R with the metafor.pdf*

```{r Vignettes, eval=FALSE, echo=TRUE}
browseVignettes("metafor")
```

## Reproducing the example

```{r example1}
dat <- escalc(measure = "RR", ai = tpos, bi = tneg, ci = cpos,
              di = cneg, data = dat.bcg, append = TRUE)
kable(dat)

res <- rma(yi, vi, data = dat)
res

forest(res, slab = paste(dat$author, dat$year, sep = ", "),
       xlim = c(-16, 6), at = log(c(0.05, 0.25, 1, 4)), atransf = exp,
       ilab = cbind(dat$tpos, dat$tneg, dat$cpos, dat$cneg),
       ilab.xpos = c(-9.5, -8, -6, -4.5), cex = 0.75)
op <- par(cex = 0.75, font = 2)
text(c(-9.5, -8, -6, -4.5), 15, c("TB+", "TB-", "TB+", "TB-"))
text(c(-8.75, -5.25), 16, c("Vaccinated", "Control"))
text(-16, 15, "Author(s) and Year", pos = 4)
text(6, 15, "Relative Risk [95% CI]", pos = 2)
par(op)
```
